---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CluReAL

<!-- badges: start -->
<!-- badges: end -->

'CluReAL' is a port of the Python implementation of the algorithm 
'CluReAL.v2', which is designed to improve an existing clustering solution by 
splitting multimodal clusters, merging akin clusters, and marking tiny 
or low-density clusters as outliers (noise). Additionally, symbolic key 
ideograms can be created to interpret clusters in high-dimensional space. 
The approach is described in detail in the article by Iglesias et al. (2021):  
<https://doi.org/10.1007/s41060-021-00275-z>. 

## Installation

<!-- You can install the released version of CluReAL from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("CluReAL") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("unmnn/CluReAL")
```
## Example

A typical CluReAL workflow:

1. Create the _cluster context_ by calling `cluster_context(x, y)`, where 
the matrix `x` represents the dataset with m observations and n dimensions, and
`y` is the integer vector of length m containing the cluster membership indices
(-1 represents noise). 
The output is a list with the following elements: 
    * `k`: number of clusters (integer)
    * `centroids`: centroid coordinates (k x n matrix)
    * `mass`: cluster sizes (integer vector of length k)
    * `mn_da`: mean Euclidean distances of the cluster members to their centroid 
  (double vector of length k)
    * `md_da`: median Euclidean distances of the cluster members to their centroid 
  (double vector of length k)
    * `sd_da`: standard deviation of the Euclidean distances of the cluster
  members to their centroid (double vector of length k)
    * `de`: Euclidean distance matrix for the centroids (k x k matrix)
    * `outliers`: number of outliers (integer)
  
2. Compute the cluster validity measures by calling `gval(cc)` using the output 
object from step 1. The output is a list with the following elements:
    * `g_str`: strict G-index (double)
    * `g_rex`: relaxed G-index (double) 
    * `g_min`: min G-index (double)
    * `oi_st`: cluster-individual strict overlap indices 
  (double vector of length k)
    * `oi_rx`: cluster-individual relaxed overlap indices 
  (double vector of length k)
    * `oi_mn`: cluster-individual min overlap indices (double vector of length k)
    * `ext_r`: extended cluster radii (double vector of length k) 
    * `str_r`: strict cluster radii (double vector of length k)
    * `vol_r`: extended-to-core ratio

3. Compute the _refinement context_ by calling 
`refinement_context(x, y, cc, gv)`.
The output is a list with the following elements:
    * `mm`: cluster multimodality flag (logical vector of length k)
    * `k_dens`: cluster-individual relative density (double vector of length k)
    * `global_c_dens`: global density (double)
    * `kinship`: cluster kinship matrix (k x k matrix): 0-itself, 
  1-parent and child, 2-relatives, 3-close friends, 4-acquaintances, 
  5-unrelated. 

4. Refine the clustering by calling `refine(x, y, cc, gv, rc)`.
The output is a list with the following elements:
    * `y`: vector of the refined cluster membership indices
    * `cc`: refined cluster context

Load all required packages:

```{r load-dependencies}
library(CluReAL)
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("ggplot2")
# install.packages("palmerpenguins")
# install.packages("patchwork")
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(patchwork)
```

We perform kmeans clustering on the Palmer penguins dataset using the variables 
`flipper_length_mm` and `bill_length_mm`. 
We [min-max normalize](https://en.wikipedia.org/wiki/Feature_scaling#Rescaling_(min-max_normalization)) the variables to unify their range.

```{r prepare}
peng <- palmerpenguins::penguins %>%
  tidyr::drop_na() %>%
  mutate(across(c(flipper_length_mm, bill_length_mm),
                ~ (.x - min(.x)) / (max(.x) - (min(.x))))) %>%
  select(flipper = flipper_length_mm, bill = bill_length_mm, species)
ggplot(peng, aes(x = flipper, y = bill, color = species)) + 
  geom_point()
```

Here, we deliberately call kmeans with a cluster count that is too high.
<!-- + we  -->
<!-- manually specify two noise points (just for demonstration purposes). -->

```{r kmeans}
set.seed(1)
clustering <- kmeans(peng[c("flipper", "bill")], centers = 6)

peng <- peng %>% mutate(cluster = as.factor(clustering$cluster))

p1 <- ggplot(peng, aes(x = flipper, y = bill, color = cluster)) + 
  geom_point() +
  labs(title = "Before refining")
p1
```

We perform the four steps of CluReAL as described above and compare the 
clustering solution before and after refining.

```{r clureal, fig.width=9, fig.height=5}
# Step 1: compute the cluster context
x <- as.matrix(peng[c("flipper", "bill")])
y <- clustering$cluster
cc <- cluster_context(x, y)
cc

# Step 2: compute the cluster validity indices
gv <- gval(cc)
gv

# Step 3: compute the refinement context
rc <- refinement_context(x, y, cc, gv)

# Step 4: refine the clustering
rf <- refine(x, y, cc, gv, rc)
rf

peng <- peng %>% mutate(c_refined = as.factor(rf$y))

p2 <- ggplot(peng, aes(x = flipper, y = bill, color = c_refined)) + 
  geom_point() +
  labs(title = "After refining")
p1 + p2
```

Draw the clustering solution ideogram before and after refining.

```{r ideogram, fig.dim=c(6,2)}
ideo_before <- draw_symbol(cc, gv, rc)

y_r <- rf$y
cc_r <- rf$cc
gv_r <- gval(cc_r)
rc_r <- refinement_context(x, y_r, cc_r, gv_r)
ideo_after <- draw_symbol(cc_r, gv_r, rc_r)

ideo_before + ideo_after
```

Interpretation:

* left: chaotic space, unreliable clusters, `g_str` < 0, `g_rex` < 0
* right: space with strong global overlap, `g_str` < 0, `g_rex` > 1

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>. -->

<!-- You can also embed plots, for example: -->

<!-- ```{r pressure, echo = FALSE} -->
<!-- plot(pressure) -->
<!-- ``` -->

<!-- In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
